---
title: "4T1 Mets RNA Seq - Association with Obesity"
author: "Andrew Chen"
date:   "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: 'united'
    toc: true
    toc_depth: 1
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r}
library(GSVA)
library(openxlsx)
library(tidyverse)
library(Biobase)
library(ComplexHeatmap)
library(circlize)
library(hypeR)

PATH <- file.path(Sys.getenv("MLAB"), "projects/brcameta/exosome/4t1_brca")
TCGAPATH <- file.path(Sys.getenv("CBM"),"TCGA-GDC/")
METABRICPATH <- file.path(Sys.getenv("CBM"), "METABRIC/")
DPATH <- file.path(Sys.getenv("CBM"),"otherStudies/RNAseq/2023-03-22-YuhanExosomeBrCa")

do_save <- FALSE
```

# Loading Data
```{r}
tcga_brca <- readRDS(file.path(TCGAPATH,"/esets_filtered/TCGA-BRCA_2020-03-22_DESeq2_log_filtered_eset.rds"))

metabric <- readRDS(file.path(METABRICPATH, "ESets/metabric_GE_ESet.rds"))
metabric_impute <- readRDS(file.path(METABRICPATH, "ESets/metabric_GE_ESet_impute.rds"))
```


```{r}
#Signatures

## Obesity signature from Fuentes-Mattei et al., JNCI 2014, Table S3.
obDiff <- read.xlsx( file.path(PATH,"data/JNCI2014_SupplementaryTable3_ObeseSignatureBrCa.xlsx") ) 
colnames(obDiff)[match(c("Gene.Title.(Definition)","Log.ratio.(Log10)","p-value*"),colnames(obDiff))] <-
    c("Gene.Title","Log10.ratio","pvalue")
obSig <- list(
    ob.up=dplyr::filter(obDiff,Log10.ratio>0) %>% select(Gene.Symbol) %>% drop_na() %>% distinct() %>% pull(),
    ob.dn=dplyr::filter(obDiff,Log10.ratio<0) %>% select(Gene.Symbol) %>% drop_na() %>% distinct() %>% pull())

diabetes_sigs <- readRDS(file.path(PATH, "data/diabetes_sigs.rds")) %>% unlist(use.names=TRUE, recursive = FALSE)

## Hallmark Signatures
hallmark_genesets <- msigdb_download(species="Homo sapiens", category="H")
hallmark_emt <- hallmark_genesets["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION"]

## Signatures from RNA-Seq experiments
all_sigs <- readRDS(file.path(PATH,"data/signatures_symbol.rds"))
all_sigs <- all_sigs[lapply(all_sigs, length) > 1]

all_sigs_human <- lapply(all_sigs, babelgene::orthologs, species="mouse", human=FALSE)
all_sigs_human <- lapply(all_sigs_human, function (x) x %>% dplyr::pull(human_symbol))

all_sigs_human <- c(all_sigs_human, obSig, diabetes_sigs, hallmark_emt)
```

# GSVA
```{r}
# GSVA
if (do_save) {
  #subtype_filter <- is.na(tcga_brca$subtype_selected)
  #tcga_brca <- tcga_brca[,!subtype_filter]
  #tcga_brca <- tcga_brca[,tcga_brca$subtype_selected == "BRCA.LumA"]
  #metabric <- metabric[,metabric$Pam50_SUBTYPE == "LumA"]
  tcga_gsva <- gsva(tcga_brca, all_sigs_human, mx.diff=TRUE, verbose=FALSE)
  metabric_gsva <- gsva(metabric, all_sigs_human, mx.diff=TRUE, verbose=FALSE)
  saveRDS(metabric_gsva, file.path(PATH, "data/metabric_gsva_res_obs.rds"))
  saveRDS(tcga_gsva, file.path(PATH, "data/tcga_gsva_res_obs.rds"))  
} else {
  tcga_gsva <- readRDS(file.path(PATH, "data/tcga_gsva_res_obs.rds"))
  metabric_gsva <- readRDS(file.path(PATH, "data/metabric_gsva_res_obs.rds"))
}
```

```{r}
# Subtype filtering
metabric_gsva <- metabric_gsva[,metabric_gsva$Pam50_SUBTYPE != "NC"]
subtype_filter <- is.na(tcga_gsva$subtype_selected)
tcga_gsva <- tcga_gsva[,!subtype_filter]

pData(tcga_gsva) <- pData(tcga_gsva) %>% mutate(pam50subtype = str_split(subtype_selected, pattern="BRCA.", simplify=TRUE)[,2])
tcga_gsva$ir_c <- t(exprs(tcga_gsva["ir_c_up",]) - exprs(tcga_gsva["ir_c_down"]))
tcga_gsva$tgfb_c <- t(exprs(tcga_gsva["tgfb_c_up",])-exprs(tcga_gsva["tgfb_c_down",]))
tcga_gsva$ir_is <- t(exprs(tcga_gsva["ir_is_up",])-exprs(tcga_gsva["ir_is_down",]))
tcga_gsva$tgfb_is <- t(exprs(tcga_gsva["tgfb_is_up",])-exprs(tcga_gsva["tgfb_is_down",]))
tcga_gsva$ob <- t(exprs(tcga_gsva["ob.up",])-exprs(tcga_gsva["ob.dn",]))

metabric_gsva$ir_c <- t(exprs(metabric_gsva["ir_c_up",]) - exprs(metabric_gsva["ir_c_down"]))
metabric_gsva$tgfb_c <- t(exprs(metabric_gsva["tgfb_c_up",])-exprs(metabric_gsva["tgfb_c_down",]))
metabric_gsva$ir_is <- t(exprs(metabric_gsva["ir_is_up",])-exprs(metabric_gsva["ir_is_down",]))
metabric_gsva$tgfb_is <- t(exprs(metabric_gsva["tgfb_is_up",])-exprs(metabric_gsva["tgfb_is_down",]))
metabric_gsva$ob <- t(exprs(metabric_gsva["ob.up",])-exprs(metabric_gsva["ob.dn",]))
```


# Boxplots
```{r}
col.pam50 <- c(LumA="pink",LumB="red",Her2="yellow",Basal="darkgray",Normal="white")
for (gset in c("ir_c", "ir_is", "tgfb_c", "ob")) {
    par(mfrow=c(1,2))
    boxplot(pData(tcga_gsva)[,gset] ~ tcga_gsva$pam50subtype, col=col.pam50, las=2,
            xlab="pam50",ylab=gset,main="TCGA" )
    boxplot(pData(metabric_gsva)[,gset] ~ metabric_gsva$Pam50_SUBTYPE, col=col.pam50, las=2,
            xlab="pam50",ylab=gset,main="METABRIC" )
}
```

# Correlation with Obesity {.tabset}

## TCGA {.tabset}
```{r}
sample_col <- hcl.colors(n=nlevels(tcga_gsva$pam50subtype %>% factor))
names(sample_col) <- levels(tcga_gsva$pam50subtype %>% factor)
ha.genes <- HeatmapAnnotation(sample_type=tcga_gsva$pam50subtype,
                              col=list(sample_type=sample_col))

png(file=file.path(PATH, "results/GSVA_tcga_obs_Heatmap.png"))
gsva_heatmap <- Heatmap(t(scale(t(exprs(tcga_gsva)))),
        name="Enrichment Score", 
        col=colorRamp2(c(-3, 0, 3), c("#072448", "white", "#ff6150")),
        top_annotation=ha.genes, 
        cluster_rows=TRUE,
        cluster_columns=TRUE,
        clustering_distance_rows="euclidean",
        clustering_method_rows="ward.D",    
        clustering_distance_columns="euclidean",
        clustering_method_columns="ward.D", 
        show_parent_dend_line=TRUE,
        column_title = "samples",
        column_dend_height = unit(5, "mm"),
        row_title="Genesets",
        row_dend_width = unit(5, "mm"),
        show_column_names=FALSE,
        show_row_names=TRUE)
draw(gsva_heatmap)
dev.off()
gsva_heatmap
```

### Corplot between Obesity/Diabetes signatures

```{r}
library(ggcorrplot)
gs_cor <- cor(t(exprs(tcga_gsva)))
```

```{r}
ob_cor <- gs_cor[grepl("ob|dz",rownames(gs_cor)),grepl("ob|dz",colnames(gs_cor))]

ggcorrplot(ob_cor, hc.order = FALSE, type = "lower",
   outline.col = "white",
   ggtheme = ggplot2::theme_gray,
   lab=TRUE,
   lab_size=3,
   colors = c("#6D9EC1", "white", "#E46726"))
```

### Exosome Signatures & Diabetes/Obesity Signatures
```{r}
cor_df <- as.data.frame(as.table(gs_cor))
colnames(cor_df) <- c("GS1", "GS2", "cor")
cor_df$GS1 <- str_replace(cor_df$GS1, pattern="dn", replacement = "down")
cor_df$GS2 <- str_replace(cor_df$GS2, pattern="dn", replacement = "down")
cor_df %>% 
  filter(grepl("ir_c|is_c|ir_is|tgfb_c|tgfb_ir", GS1)) %>%
  filter(grepl("dz|ob", GS2)) %>%
  separate(col = GS2, into = c("GS2", "direction"), sep = "\\.") %>% 
  ggplot() + 
  geom_boxplot(aes(x=GS1, y=cor, fill=direction)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Correlation between DE signatures and diabetes/obesity signatures")
```

### Up-Down
```{r}
tcga_updown_cor <- cor(pData(tcga_gsva)[,c("ir_c", "ir_is", "tgfb_c", "tgfb_is", "ob")])

ggcorrplot(tcga_updown_cor, hc.order = FALSE, type = "lower",
   outline.col = "white",
   ggtheme = ggplot2::theme_gray,
   lab=TRUE,
   lab_size=3,
   colors = c("#6D9EC1", "white", "#E46726"))
```


## Metabric {.tabset}
```{r}
sample_col <- hcl.colors(n=nlevels(metabric_gsva$Pam50_SUBTYPE %>% factor))
names(sample_col) <- levels(metabric_gsva$Pam50_SUBTYPE %>% factor)
ha.genes <- HeatmapAnnotation(sample_type=metabric_gsva$Pam50_SUBTYPE,
                              col=list(sample_type=sample_col))

png(file=file.path(PATH, "results/GSVA_metabric_obs_Heatmap.png"))
gsva_heatmap <- Heatmap(t(scale(t(exprs(metabric_gsva)))),
        name="Enrichment Score", 
        col=colorRamp2(c(-3, 0, 3), c("#072448", "white", "#ff6150")),
        top_annotation=ha.genes, 
        cluster_rows=TRUE,
        cluster_columns=TRUE,
        clustering_distance_rows="euclidean",
        clustering_method_rows="ward.D",    
        clustering_distance_columns="euclidean",
        clustering_method_columns="ward.D", 
        show_parent_dend_line=TRUE,
        column_title = "samples",
        column_dend_height = unit(5, "mm"),
        row_title="Genesets",
        row_dend_width = unit(5, "mm"),
        show_column_names=FALSE,
        show_row_names=TRUE)
draw(gsva_heatmap)
dev.off()
gsva_heatmap
```

### Corplot between Obesity/Diabetes signatures
```{r}
gs_cor <- cor(t(exprs(metabric_gsva)))

ob_cor <- gs_cor[grepl("ob|dz",rownames(gs_cor)),grepl("ob|dz",colnames(gs_cor))]

ggcorrplot(ob_cor, hc.order = FALSE, type = "lower",
   outline.col = "white",
   ggtheme = ggplot2::theme_gray,
   lab=TRUE,
   lab_size=3,
   colors = c("#6D9EC1", "white", "#E46726"))
```

### Exosome Signatures & Diabetes/Obesity Signatures

```{r}
cor_df <- as.data.frame(as.table(gs_cor))
colnames(cor_df) <- c("GS1", "GS2", "cor")
cor_df$GS1 <- str_replace(cor_df$GS1, pattern="dn", replacement = "down")
cor_df$GS2 <- str_replace(cor_df$GS2, pattern="dn", replacement = "down")
cor_df %>% 
  filter(grepl("ir_c|is_c|ir_is|tgfb_c|tgfb_ir", GS1)) %>%
  filter(grepl("dz|ob", GS2)) %>%
  separate(col = GS2, into = c("GS2", "direction"), sep = "\\.") %>% 
  ggplot() + 
  geom_boxplot(aes(x=GS1, y=cor, fill=direction)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Up-Down
```{r}
metabric_updown_cor <- cor(pData(metabric_gsva)[,c("ir_c", "ir_is", "tgfb_c", "tgfb_is", "ob")])

ggcorrplot(metabric_updown_cor, hc.order = FALSE, type = "lower",
   outline.col = "white",
   ggtheme = ggplot2::theme_gray,
   lab=TRUE,
   lab_size=3,
   colors = c("#6D9EC1", "white", "#E46726"))
```

## Similarity between Metabric and TCGA
```{r}
updown_cor_combined <- tcga_updown_cor
updown_cor_combined[upper.tri(updown_cor_combined)] <- metabric_updown_cor[upper.tri(metabric_updown_cor)]
diag(updown_cor_combined) <- NA
ggcorrplot(updown_cor_combined,
           method="square",
           type="full",
           outline.col = "white",
           ggtheme = ggplot2::theme_gray,
           lab=TRUE,
           lab_size=3,
           colors = c("#6D9EC1", "white", "#E46726"))
```


# KS-based enrichment {.tabset}
```{r}
## Add gene symbols to DESeq tables
dat <- readRDS(file.path(PATH, "data/4T1_mets_sExp.rds"))
deseq_list <- readRDS(file.path(PATH,"data/deseq_list.rds"))
hallmark_genesets <- msigdb_download(species="Mus musculus", category="H")
hallmark_genesets$OBESITY <- obSig

deseq_list <- lapply(deseq_list, function(Z) {
  as.data.frame(Z) %>% 
    tibble::rownames_to_column(var="ensembl_id") %>%
    dplyr::mutate(geneID=unlist(rowData(dat)[ensembl_id,"mgi_symbol"]))
  })
```

```{r}
## rank by up-regulation
rank_signatures_up <- lapply(
  deseq_list, function(sig) sig %>% 
    dplyr::filter(geneID!="") %>%
    dplyr::arrange(desc(log2FoldChange)) %>%
    dplyr::select(geneID,log2FoldChange) %>%
    tibble::deframe())
names(rank_signatures_up) <- paste(names(rank_signatures_up),"up",sep="_")

## rank by down-regulation
rank_signatures_dn <- lapply(
  deseq_list,function(sig) sig %>% 
    dplyr::filter(geneID!="") %>%
    dplyr::arrange(log2FoldChange) %>%
    dplyr::select(geneID,log2FoldChange) %>%
    tibble::deframe())
names(rank_signatures_dn) <- paste(names(rank_signatures_dn),"dn",sep="_")

rank_signatures <- c(rank_signatures_up,rank_signatures_dn)
rank_signatures <- rank_signatures[order(names(rank_signatures),decreasing = TRUE)]
```

### Hallmarks + Pamm

```{r}
#all_genesets <- c(hallmark_genesets, pamm_genelists)
max_fdr <- 0.05
ks_hall <- hypeR::hypeR(signature=rank_signatures,genesets=hallmark_genesets,test="ks",fdr=max_fdr,plotting=TRUE)
hyp_dots(ks_hall,merge=TRUE,fdr=max_fdr/5,top=25) + ggtitle(paste("FDR ≤", max_fdr/5))
```

```{r}
#hypeR::rctbl_build(ks_hall, show_hmaps=FALSE)
```

## IS C UP
```{r}
print(ks_hall$data$is_c_up$plots[ks_hall$data$is_c_up$data$fdr<=0.01])
```

## IS C Down
```{r}
print(ks_hall$data$is_c_dn$plots[ks_hall$data$is_c_dn$data$fdr<=0.01])
```

## IR IS Up
```{r}
print(ks_hall$data$ir_is_up$plots[ks_hall$data$ir_is_up$data$fdr<=0.01])
```

## IR IS Down
```{r}
print(ks_hall$data$ir_is_dn$plots[ks_hall$data$ir_is_dn$data$fdr<=0.01])
```

## IR C Up
```{r}
print(ks_hall$data$ir_c_up$plots[ks_hall$data$ir_c_up$data$fdr<=0.01])
```

## IR C Down
```{r}
print(ks_hall$data$ir_c_dn$plots[ks_hall$data$ir_c_dn$data$fdr<=0.01])
```
