---
title: "Recontextualized 4T1 RNA Seq - Survival Analysis based on GSVA Projections"
author: "Andrew Chen"
date:   "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: 'united'
    toc: true
    toc_depth: 1
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r}
library(Biobase)
library(survminer)
library(survival)
library(tidyverse)
library(DT)
theme_set(theme_gray(base_size = 25))
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = FALSE,
                      fig.height = 6,
                      fig.width = 7, 
                      cache = TRUE)

PATH <- file.path(Sys.getenv("MLAB"), "projects/brcameta/exosome/4t1_brca")
TCGAPATH <- file.path(Sys.getenv("CBM"), "TCGA-GDC/")
DPATH <- file.path(Sys.getenv("CBM"), "otherStudies/RNAseq/2023-03-22-YuhanExosomeBrCa")

do_save <- FALSE
```

# Loading Data
```{r}
tcga_gsva <- readRDS(file.path(PATH, "data/tcga_gsva_recon.rds"))
metabric_gsva <- readRDS(file.path(PATH, "data/metabric_gsva_recon.rds"))

# Signatures

all_sigs <- readRDS(file.path(PATH, "data/signatures_symbol.rds"))
all_sigs <- all_sigs[lapply(all_sigs, length) > 1]

all_sigs_human <- lapply(all_sigs, babelgene::orthologs, species = "mouse", human = FALSE)
all_sigs_human <- lapply(all_sigs_human, function(x) x %>% dplyr::pull(human_symbol))
```

Data Filtering
```{r}
na_filter <- !is.na(tcga_gsva$vital_status)
missing_death_day_filter <- !((tcga_gsva$vital_status == "Dead") & is.na(tcga_gsva$days_to_death))
subtype_filter <- !is.na(tcga_gsva$subtype_selected)
data_filter <- na_filter & missing_death_day_filter & subtype_filter

tcga_gsva <- tcga_gsva[, data_filter]
```

> Censoring: If samples are alive, then they lived at least as long as the day to their last followup. For 5 year threshold, those who died after 5 years, should be right-censored, and labeled to have lived at least as long as 'days to death' + 1.

```{r}
pData(tcga_gsva) <- pData(tcga_gsva) %>%
  rownames_to_column("ID") %>%
  mutate(time = if_else(!is.na(days_to_death), days_to_death, days_to_last_follow_up)) %>%
  mutate(time_5 = if_else(as.numeric(time) < 1825.0, as.numeric(time), 1826.0)) %>%
  mutate(vital_status_1 = if_else(vital_status == "Alive", 1, 2)) %>%
  mutate(vital_status_5 = if_else(vital_status == "Dead" & (time_5 > 1825.0), 1, vital_status_1)) %>%
  column_to_rownames("ID")

pData(metabric_gsva) <- pData(metabric_gsva) %>%
  mutate(time = OS_MONTHS * 30.437) %>%
  mutate(time_5 = if_else(as.numeric(time) < 1825.0, as.numeric(time), 1826.0)) %>%
  mutate(vital_status_1 = if_else(OS_STATUS == "LIVING", 1, 2)) %>%
  mutate(vital_status_5 = if_else(OS_STATUS == "DECEASED" & (time_5 > 1825.0), 1, vital_status_1))
```

# Survfit All subtypes {.tabset}

```{r}
# Finding Median Thresholds

tcga_threshold_ir_c <- median(exprs(tcga_gsva["ir_c_up", ]) - exprs(tcga_gsva["ir_c_down"]))
tcga_threshold_is_c <- median(exprs(tcga_gsva["is_c_up", ]) - exprs(tcga_gsva["is_c_down", ]))
tcga_threshold_ir_is <- median(exprs(tcga_gsva["ir_is_up", ]) - exprs(tcga_gsva["ir_is_down", ]))
tcga_threshold_tgfb_c <- median(exprs(tcga_gsva["tgfb_c_up", ]) - exprs(tcga_gsva["tgfb_c_down", ]))

metabric_threshold_ir_c <- median(exprs(metabric_gsva["ir_c_up", ]) - exprs(metabric_gsva["ir_c_down"]))
metabric_threshold_is_c <- median(exprs(metabric_gsva["is_c_up", ]) - exprs(metabric_gsva["is_c_down", ]))
metabric_threshold_ir_is <- median(exprs(metabric_gsva["ir_is_up", ]) - exprs(metabric_gsva["ir_is_down", ]))
metabric_threshold_tgfb_c <- median(exprs(metabric_gsva["tgfb_c_up", ]) - exprs(metabric_gsva["tgfb_c_down", ]))

tcga_threshold_ir_c_up <- median(exprs(tcga_gsva["ir_c_up", ]))
tcga_threshold_is_c_up <- median(exprs(tcga_gsva["is_c_up", ]))
tcga_threshold_ir_is_up <- median(exprs(tcga_gsva["ir_is_up", ]))
tcga_threshold_tgfb_c_up <- median(exprs(tcga_gsva["tgfb_c_up", ]))

metabric_threshold_ir_c_up <- median(exprs(metabric_gsva["ir_c_up", ]))
metabric_threshold_is_c_up <- median(exprs(metabric_gsva["is_c_up", ]))
metabric_threshold_ir_is_up <- median(exprs(metabric_gsva["ir_is_up", ]))
metabric_threshold_tgfb_c_up <- median(exprs(metabric_gsva["tgfb_c_up", ]))
```

Adding GSVA data
```{r}
tcga_gsva$ir_c <- t(exprs(tcga_gsva["ir_c_up", ]) - exprs(tcga_gsva["ir_c_down"]))
tcga_gsva$is_c <- t(exprs(tcga_gsva["is_c_up", ]) - exprs(tcga_gsva["is_c_down", ]))
tcga_gsva$ir_is <- t(exprs(tcga_gsva["ir_is_up", ]) - exprs(tcga_gsva["ir_is_down", ]))
tcga_gsva$tgfb_c <- t(exprs(tcga_gsva["tgfb_c_up", ]) - exprs(tcga_gsva["tgfb_c_down", ]))

tcga_gsva$ir_c_stat <- with(tcga_gsva, ifelse(tcga_gsva$ir_c <= tcga_threshold_ir_c, "low", "high"))
tcga_gsva$is_c_stat <- with(tcga_gsva, ifelse(tcga_gsva$is_c <= tcga_threshold_is_c, "low", "high"))
tcga_gsva$ir_is_stat <- with(tcga_gsva, ifelse(tcga_gsva$ir_is <= tcga_threshold_ir_is, "low", "high"))
tcga_gsva$tgfb_c_stat <- with(tcga_gsva, ifelse(tcga_gsva$tgfb_c <= tcga_threshold_tgfb_c, "low", "high"))

tcga_gsva$ir_c_up <- t(exprs(tcga_gsva["ir_c_up", ]))
tcga_gsva$is_c_up <- t(exprs(tcga_gsva["is_c_up", ]))
tcga_gsva$ir_is_up <- t(exprs(tcga_gsva["ir_is_up", ]))
tcga_gsva$tgfb_c_up <- t(exprs(tcga_gsva["tgfb_c_up", ]))
tcga_gsva$ir_c_up_stat <- with(tcga_gsva, ifelse(tcga_gsva$ir_c_up <= tcga_threshold_ir_c_up, "low", "high"))
tcga_gsva$is_c_up_stat <- with(tcga_gsva, ifelse(tcga_gsva$is_c_up <= tcga_threshold_is_c_up, "low", "high"))
tcga_gsva$ir_is_up_stat <- with(tcga_gsva, ifelse(tcga_gsva$ir_is_up <= tcga_threshold_ir_is_up, "low", "high"))
tcga_gsva$tgfb_c_up_stat <- with(tcga_gsva, ifelse(tcga_gsva$tgfb_c_up <= tcga_threshold_tgfb_c_up, "low", "high"))

metabric_gsva$ir_c <- t(exprs(metabric_gsva["ir_c_up", ]) - exprs(metabric_gsva["ir_c_down"]))
metabric_gsva$is_c <- t(exprs(metabric_gsva["is_c_up", ]) - exprs(metabric_gsva["is_c_down", ]))
metabric_gsva$ir_is <- t(exprs(metabric_gsva["ir_is_up", ]) - exprs(metabric_gsva["ir_is_down", ]))
metabric_gsva$tgfb_c <- t(exprs(metabric_gsva["tgfb_c_up", ]) - exprs(metabric_gsva["tgfb_c_down", ]))
metabric_gsva$ir_c_stat <- with(metabric_gsva, ifelse(metabric_gsva$ir_c <= metabric_threshold_ir_c, "low", "high"))
metabric_gsva$is_c_stat <- with(metabric_gsva, ifelse(metabric_gsva$is_c <= metabric_threshold_is_c, "low", "high"))
metabric_gsva$ir_is_stat <- with(metabric_gsva, ifelse(metabric_gsva$ir_is <= metabric_threshold_ir_is, "low", "high"))
metabric_gsva$tgfb_c_stat <- with(tcga_gsva, ifelse(metabric_gsva$tgfb_c <= metabric_threshold_tgfb_c, "low", "high"))

metabric_gsva$ir_c_up <- t(exprs(metabric_gsva["ir_c_up", ]))
metabric_gsva$is_c_up <- t(exprs(metabric_gsva["is_c_up", ]))
metabric_gsva$ir_is_up <- t(exprs(metabric_gsva["ir_is_up", ]))
metabric_gsva$tgfb_c_up <- t(exprs(metabric_gsva["tgfb_c_up", ]))
metabric_gsva$ir_c_up_stat <- with(metabric_gsva, ifelse(metabric_gsva$ir_c_up <= metabric_threshold_ir_c_up, "low", "high"))
metabric_gsva$is_c_up_stat <- with(metabric_gsva, ifelse(metabric_gsva$is_c_up <= metabric_threshold_is_c_up, "low", "high"))
metabric_gsva$ir_is_up_stat <- with(metabric_gsva, ifelse(metabric_gsva$ir_is_up <= metabric_threshold_ir_is_up, "low", "high"))
metabric_gsva$tgfb_c_up_stat <- with(metabric_gsva, ifelse(metabric_gsva$tgfb_c_up <= metabric_threshold_tgfb_c_up, "low", "high"))
```

## IR_C_UP
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time), tcga_gsva$vital_status_1) ~ tcga_gsva$ir_c_up_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time_5), tcga_gsva$vital_status_5) ~ tcga_gsva$ir_c_up_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA w/ 5-year threshold"
)
```
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time), metabric_gsva$vital_status_1) ~ metabric_gsva$ir_c_up_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time_5), metabric_gsva$vital_status_5) ~ metabric_gsva$ir_c_up_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC w/ 5-year threshold"
)
```

## IR_C
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time), tcga_gsva$vital_status_1) ~ tcga_gsva$ir_c_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time_5), tcga_gsva$vital_status_5) ~ tcga_gsva$ir_c_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA w/ 5-year threshold"
)
```
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time), metabric_gsva$vital_status_1) ~ metabric_gsva$ir_c_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time_5), metabric_gsva$vital_status_5) ~ metabric_gsva$ir_c_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC w/ 5-year threshold"
)
```

## IS_C_UP
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time), tcga_gsva$vital_status_1) ~ tcga_gsva$is_c_up_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time_5), tcga_gsva$vital_status_5) ~ tcga_gsva$is_c_up_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA w/ 5-year threshold"
)
```
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time), metabric_gsva$vital_status_1) ~ metabric_gsva$is_c_up_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time_5), metabric_gsva$vital_status_5) ~ metabric_gsva$is_c_up_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC w/ 5-year threshold"
)
```

## IS_C
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time), tcga_gsva$vital_status_1) ~ tcga_gsva$is_c_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time_5), tcga_gsva$vital_status_5) ~ tcga_gsva$is_c_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA w/ 5-year threshold"
)
```
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time), metabric_gsva$vital_status_1) ~ metabric_gsva$is_c_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time_5), metabric_gsva$vital_status_5) ~ metabric_gsva$is_c_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC w/ 5-year threshold"
)
```

## IR_IS_UP
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time), tcga_gsva$vital_status_1) ~ tcga_gsva$ir_is_up_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time_5), tcga_gsva$vital_status_5) ~ tcga_gsva$ir_is_up_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA w/ 5-year threshold"
)
```
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time), metabric_gsva$vital_status_1) ~ metabric_gsva$ir_is_up_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time_5), metabric_gsva$vital_status_5) ~ metabric_gsva$ir_is_up_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC w/ 5-year threshold"
)
```

## IR_IS
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time), tcga_gsva$vital_status_1) ~ tcga_gsva$ir_is_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time_5), tcga_gsva$vital_status_5) ~ tcga_gsva$ir_is_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA w/ 5-year threshold"
)
```
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time), metabric_gsva$vital_status_1) ~ metabric_gsva$ir_is_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time_5), metabric_gsva$vital_status_5) ~ metabric_gsva$ir_is_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC w/ 5-year threshold"
)
```

## TGFB_C_UP
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time), tcga_gsva$vital_status_1) ~ tcga_gsva$tgfb_c_up_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time_5), tcga_gsva$vital_status_5) ~ tcga_gsva$tgfb_c_up_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA w/ 5-year threshold"
)
```
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time), metabric_gsva$vital_status_1) ~ metabric_gsva$tgfb_c_up_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time_5), metabric_gsva$vital_status_5) ~ metabric_gsva$tgfb_c_up_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC w/ 5-year threshold"
)
```

## TGFB_C
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time), tcga_gsva$vital_status_1) ~ tcga_gsva$tgfb_c_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva$time_5), tcga_gsva$vital_status_5) ~ tcga_gsva$tgfb_c_stat),
  data = tcga_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA w/ 5-year threshold"
)
```
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time), metabric_gsva$vital_status_1) ~ metabric_gsva$tgfb_c_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva$time_5), metabric_gsva$vital_status_5) ~ metabric_gsva$tgfb_c_stat),
  data = metabric_gsva,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC w/ 5-year threshold"
)
```


# Survfit Basal {.tabset}

```{r}
# Thresholds
tcga_basal_filter <- tcga_gsva$subtype_selected == "BRCA.Basal"
tcga_gsva_basal <- tcga_gsva[, tcga_basal_filter]
metabric_basal_filter <- metabric_gsva$Pam50_SUBTYPE == "Basal"
metabric_gsva_basal <- metabric_gsva[, metabric_basal_filter]

tcga_threshold_ir_c <- median(exprs(tcga_gsva_basal["ir_c_up", ]) - exprs(tcga_gsva_basal["ir_c_down", ]))
tcga_threshold_is_c <- median(exprs(tcga_gsva_basal["is_c_up", ]) - exprs(tcga_gsva_basal["is_c_down", ]))
tcga_threshold_ir_is <- median(exprs(tcga_gsva_basal["ir_is_up", ]) - exprs(tcga_gsva_basal["ir_is_down", ]))
tcga_threshold_tgfb_c <- median(exprs(tcga_gsva_basal["tgfb_c_up", ]) - exprs(tcga_gsva_basal["tgfb_c_down", ]))
metabric_threshold_ir_c <- median(exprs(metabric_gsva_basal["ir_c_up", ]) - exprs(metabric_gsva_basal["ir_c_down", ]))
metabric_threshold_is_c <- median(exprs(metabric_gsva_basal["is_c_up", ]) - exprs(metabric_gsva_basal["is_c_down", ]))
metabric_threshold_ir_is <- median(exprs(metabric_gsva_basal["ir_is_up", ]) - exprs(metabric_gsva_basal["ir_is_down", ]))
metabric_threshold_tgfb_c <- median(exprs(metabric_gsva_basal["tgfb_c_up", ]) - exprs(metabric_gsva_basal["tgfb_c_down", ]))

tcga_threshold_ir_c_up <- median(exprs(tcga_gsva_basal["ir_c_up", ]))
tcga_threshold_is_c_up <- median(exprs(tcga_gsva_basal["is_c_up", ]))
tcga_threshold_ir_is_up <- median(exprs(tcga_gsva_basal["ir_is_up", ]))
tcga_threshold_tgfb_c_up <- median(exprs(tcga_gsva_basal["tgfb_c_up", ]))
metabric_threshold_ir_c_up <- median(exprs(metabric_gsva_basal["ir_c_up", ]))
metabric_threshold_is_c_up <- median(exprs(metabric_gsva_basal["is_c_up", ]))
metabric_threshold_ir_is_up <- median(exprs(metabric_gsva_basal["ir_is_up", ]))
metabric_threshold_tgfb_c_up <- median(exprs(metabric_gsva_basal["tgfb_c_up", ]))
```

Adding GSVA data
```{r}
tcga_gsva_basal$ir_c <- t(exprs(tcga_gsva_basal["ir_c_up", ]) - exprs(tcga_gsva_basal["ir_c_down", ]))
tcga_gsva_basal$is_c <- t(exprs(tcga_gsva_basal["is_c_up", ]) - exprs(tcga_gsva_basal["is_c_down", ]))
tcga_gsva_basal$ir_is <- t(exprs(tcga_gsva_basal["ir_is_up", ]) - exprs(tcga_gsva_basal["ir_is_down", ]))
tcga_gsva_basal$tgfb_c <- t(exprs(tcga_gsva_basal["tgfb_c_up", ]) - exprs(tcga_gsva_basal["tgfb_c_down", ]))
metabric_gsva_basal$ir_c <- t(exprs(metabric_gsva_basal["ir_c_up", ]) - exprs(metabric_gsva_basal["ir_c_down", ]))
metabric_gsva_basal$is_c <- t(exprs(metabric_gsva_basal["is_c_up", ]) - exprs(metabric_gsva_basal["is_c_down", ]))
metabric_gsva_basal$ir_is <- t(exprs(metabric_gsva_basal["ir_is_up", ]) - exprs(metabric_gsva_basal["ir_is_down", ]))
metabric_gsva_basal$tgfb_c <- t(exprs(metabric_gsva_basal["tgfb_c_up", ]) - exprs(metabric_gsva_basal["tgfb_c_down", ]))

tcga_gsva_basal$ir_c_up <- t(exprs(tcga_gsva_basal["ir_c_up", ]))
tcga_gsva_basal$is_c_up <- t(exprs(tcga_gsva_basal["is_c_up", ]))
tcga_gsva_basal$ir_is_up <- t(exprs(tcga_gsva_basal["ir_is_up", ]))
tcga_gsva_basal$tgfb_c_up <- t(exprs(tcga_gsva_basal["tgfb_c_up", ]))
metabric_gsva_basal$ir_c_up <- t(exprs(metabric_gsva_basal["ir_c_up", ]))
metabric_gsva_basal$is_c_up <- t(exprs(metabric_gsva_basal["is_c_up", ]))
metabric_gsva_basal$ir_is_up <- t(exprs(metabric_gsva_basal["ir_is_up", ]))
metabric_gsva_basal$tgfb_c_up <- t(exprs(metabric_gsva_basal["tgfb_c_up", ]))

tcga_gsva_basal$ir_c_stat <- with(tcga_gsva_basal, ifelse(tcga_gsva_basal$ir_c <= tcga_threshold_ir_c, "low", "high"))
tcga_gsva_basal$is_c_stat <- with(tcga_gsva_basal, ifelse(tcga_gsva_basal$is_c <= tcga_threshold_is_c, "low", "high"))
tcga_gsva_basal$ir_is_stat <- with(tcga_gsva_basal, ifelse(tcga_gsva_basal$ir_is <= tcga_threshold_ir_is, "low", "high"))
tcga_gsva_basal$tgfb_c_stat <- with(tcga_gsva_basal, ifelse(tcga_gsva_basal$tgfb_c <= tcga_threshold_tgfb_c, "low", "high"))
tcga_gsva_basal$ir_c_up_stat <- with(tcga_gsva_basal, ifelse(tcga_gsva_basal$ir_c_up <= tcga_threshold_ir_c_up, "low", "high"))
tcga_gsva_basal$is_c_up_stat <- with(tcga_gsva_basal, ifelse(tcga_gsva_basal$is_c_up <= tcga_threshold_is_c_up, "low", "high"))
tcga_gsva_basal$ir_is_up_stat <- with(tcga_gsva_basal, ifelse(tcga_gsva_basal$ir_is_up <= tcga_threshold_ir_is_up, "low", "high"))
tcga_gsva_basal$tgfb_c_up_stat <- with(tcga_gsva_basal, ifelse(tcga_gsva_basal$tgfb_c_up <= tcga_threshold_tgfb_c_up, "low", "high"))

metabric_gsva_basal$ir_c_stat <- with(metabric_gsva_basal, ifelse(metabric_gsva_basal$ir_c <= metabric_threshold_ir_c, "low", "high"))
metabric_gsva_basal$is_c_stat <- with(metabric_gsva_basal, ifelse(metabric_gsva_basal$is_c <= metabric_threshold_is_c, "low", "high"))
metabric_gsva_basal$ir_is_stat <- with(metabric_gsva_basal, ifelse(metabric_gsva_basal$ir_is <= metabric_threshold_ir_is, "low", "high"))
metabric_gsva_basal$tgfb_c_stat <- with(metabric_gsva_basal, ifelse(metabric_gsva_basal$tgfb_c <= metabric_threshold_tgfb_c, "low", "high"))
metabric_gsva_basal$ir_c_up_stat <- with(metabric_gsva_basal, ifelse(metabric_gsva_basal$ir_c_up <= metabric_threshold_ir_c_up, "low", "high"))
metabric_gsva_basal$is_c_up_stat <- with(metabric_gsva_basal, ifelse(metabric_gsva_basal$is_c_up <= metabric_threshold_is_c_up, "low", "high"))
metabric_gsva_basal$ir_is_up_stat <- with(metabric_gsva_basal, ifelse(metabric_gsva_basal$ir_is_up <= metabric_threshold_ir_is_up, "low", "high"))
metabric_gsva_basal$tgfb_c_up_stat <- with(metabric_gsva_basal, ifelse(metabric_gsva_basal$tgfb_c_up <= metabric_threshold_tgfb_c_up, "low", "high"))
```

## IR_C_UP
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time), tcga_gsva_basal$vital_status_1) ~ tcga_gsva_basal$ir_c_up_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time_5), tcga_gsva_basal$vital_status_5) ~ tcga_gsva_basal$ir_c_up_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA w/ 5-year threshold"
)
```
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time), metabric_gsva_basal$vital_status_1) ~ metabric_gsva_basal$ir_c_up_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "Metabric"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time_5), metabric_gsva_basal$vital_status_5) ~ metabric_gsva_basal$ir_c_up_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "Metabric w/ 5-year threshold"
)
```

## IR_C
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time), tcga_gsva_basal$vital_status_1) ~ tcga_gsva_basal$ir_c_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time_5), tcga_gsva_basal$vital_status_5) ~ tcga_gsva_basal$ir_c_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA w/ 5-year threshold"
)
```
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time), metabric_gsva_basal$vital_status_1) ~ metabric_gsva_basal$ir_c_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "Metabric"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time_5), metabric_gsva_basal$vital_status_5) ~ metabric_gsva_basal$ir_c_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "Metabric w/ 5-year threshold"
)
```

## IS_C_UP
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time), tcga_gsva_basal$vital_status_1) ~ tcga_gsva_basal$is_c_up_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time_5), tcga_gsva_basal$vital_status_5) ~ tcga_gsva_basal$is_c_up_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "w/ 5-year threshold"
)
```

```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time), metabric_gsva_basal$vital_status_1) ~ metabric_gsva_basal$is_c_up_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "Metabric"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time_5), metabric_gsva_basal$vital_status_5) ~ metabric_gsva_basal$is_c_up_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "Metabric w/ 5-year threshold"
)
```

## IS_C
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time), tcga_gsva_basal$vital_status_1) ~ tcga_gsva_basal$is_c_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time_5), tcga_gsva_basal$vital_status_5) ~ tcga_gsva_basal$is_c_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "w/ 5-year threshold"
)
```

```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time), metabric_gsva_basal$vital_status_1) ~ metabric_gsva_basal$is_c_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "Metabric"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time_5), metabric_gsva_basal$vital_status_5) ~ metabric_gsva_basal$is_c_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "Metabric w/ 5-year threshold"
)
```

## IR_IS_UP
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time), tcga_gsva_basal$vital_status_1) ~ tcga_gsva_basal$ir_is_up_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time_5), tcga_gsva_basal$vital_status_5) ~ tcga_gsva_basal$ir_is_up_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA w/ 5-year threshold"
)
```

```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time), metabric_gsva_basal$vital_status_1) ~ metabric_gsva_basal$ir_is_up_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "Metabric"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time_5), metabric_gsva_basal$vital_status_5) ~ metabric_gsva_basal$ir_is_up_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "Metabric w/ 5-year threshold"
)
```

## IR_IS
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time), tcga_gsva_basal$vital_status_1) ~ tcga_gsva_basal$ir_is_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time_5), tcga_gsva_basal$vital_status_5) ~ tcga_gsva_basal$ir_is_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA w/ 5-year threshold"
)
```

```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time), metabric_gsva_basal$vital_status_1) ~ metabric_gsva_basal$ir_is_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "Metabric"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time_5), metabric_gsva_basal$vital_status_5) ~ metabric_gsva_basal$ir_is_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "Metabric w/ 5-year threshold",
  ggtheme =   theme(axis.title = element_text(size=10),
                    axis.title.x = element_text(size=15),
                    axis.title.y = element_text(size=15),
                    legend.key.size = unit(1, 'cm'),
                    legend.title = element_text(size=15),                                       
                    legend.text = element_text(size=15)) 
) 
```

## TGFB_C_UP
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time), tcga_gsva_basal$vital_status_1) ~ tcga_gsva_basal$tgfb_c_up_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time_5), tcga_gsva_basal$vital_status_5) ~ tcga_gsva_basal$tgfb_c_up_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA w/ 5-year threshold"
)
```
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time), metabric_gsva_basal$vital_status_1) ~ metabric_gsva_basal$tgfb_c_up_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time_5), metabric_gsva_basal$vital_status_5) ~ metabric_gsva_basal$tgfb_c_up_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC w/ 5-year threshold"
)
```

## TGFB_C
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time), tcga_gsva_basal$vital_status_1) ~ tcga_gsva_basal$tgfb_c_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(tcga_gsva_basal$time_5), tcga_gsva_basal$vital_status_5) ~ tcga_gsva_basal$tgfb_c_stat),
  data = tcga_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "TCGA w/ 5-year threshold"
)
```
```{r}
ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time), metabric_gsva_basal$vital_status_1) ~ metabric_gsva_basal$tgfb_c_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC"
)

ggsurvplot(
  fit = survfit(Surv(as.numeric(metabric_gsva_basal$time_5), metabric_gsva_basal$vital_status_5) ~ metabric_gsva_basal$tgfb_c_stat),
  data = metabric_gsva_basal,
  xlab = "Days",
  ylab = "Overall survival probability",
  conf.int = TRUE,
  pval = TRUE,
  title = "METABRIC w/ 5-year threshold"
)
```

# Cox Proportional Hazard

TCGA
```{r}
pData(tcga_gsva)$subtype_selected <- relevel(factor(pData(tcga_gsva)$subtype_selected), "BRCA.Normal")
tcga.cox1 <- coxph(Surv(as.numeric(time), vital_status_1) ~ subtype_selected, data = pData(tcga_gsva))
tcga.cox2 <- coxph(Surv(as.numeric(time), vital_status_1) ~ ir_is + subtype_selected, data = pData(tcga_gsva))
tcga.cox3 <- coxph(Surv(as.numeric(time), vital_status_1) ~ ir_is + subtype_selected + age_at_index, data = pData(tcga_gsva))
anova(tcga.cox1, tcga.cox2, tcga.cox3)
```

```{r}
summary(tcga.cox3)
```


Metabric
```{r}
pData(metabric_gsva)$Pam50_SUBTYPE <- relevel(factor(pData(metabric_gsva)$Pam50_SUBTYPE), "Normal")
metabric.cox1 <- coxph(Surv(as.numeric(time), vital_status_1) ~ Pam50_SUBTYPE, data = pData(metabric_gsva))
metabric.cox2 <- coxph(Surv(as.numeric(time), vital_status_1) ~ ir_is + Pam50_SUBTYPE, data = pData(metabric_gsva))
metabric.cox3 <- coxph(Surv(as.numeric(time), vital_status_1) ~ ir_is + Pam50_SUBTYPE + AGE_AT_DIAGNOSIS, data = pData(metabric_gsva))
anova(metabric.cox1, metabric.cox2, metabric.cox3)
```
```{r}
summary(metabric.cox3)
```
