---
title: "Yuhan Primary Tumor RNA Seq - EDA"
author: "Andrew Chen"
date:   "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: 'united'
    toc: true
    toc_depth: 1
    toc_float: true
    df_print: paged
    code_folding: hide
    fig_width: 6
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r settings}
library(reshape2)
library(pheatmap)
library(mclust)
library(SummarizedExperiment)
library(DESeq2)
library(tidyverse)
library(readxl)

PATH <- file.path(Sys.getenv("MLAB"),"projects/brcameta/projects/exosome/4t1_primary_tumor/")
DPATH <- file.path(Sys.getenv("CBM"),"otherStudies/RNAseq/")
source(file.path(Sys.getenv("CBMGIT"),"MLscripts/R/rm_low_rnaseq_counts.R"))

do_save <- TRUE
```

# Loading Data
```{r}
invitro_se <- readRDS(file.path(DPATH, "2024-10-15_Yuhan_miRNA_4T1/processed/star_salmon/salmon.merged.gene_counts.rds"))

invivo_se <- readRDS(file.path(DPATH, "2024-10-15_Yuhan_PrimaryTumor/processed/star_salmon/salmon.merged.gene_counts.rds"))
```

## Gene Filtering
```{r}
# Removing genes with less than 1 read per million across 3 samples.
# Over 30K genes are removed in this step.
invitro.filtered <- rm_low_rnaseq_counts(invitro_se, min_samples = 3)
invivo.filtered <- rm_low_rnaseq_counts(invivo_se, min_samples = 3)

invitro.filtered@colData[invitro.filtered@colData$files == "miR1514D_4", "files"] <- "miR1454D_2"
rownames(invitro.filtered@colData) <- invitro.filtered@colData$files
colnames(invitro.filtered) <- colnames(invitro.filtered@assays@data$counts) <- colnames(invitro.filtered@assays@data$abundance) <- rownames(invitro.filtered@colData)
```

## Redefining ColData
```{r}
invitro.filtered@colData <- invitro.filtered@colData %>% 
  as.data.frame %>% 
  dplyr::select(files) %>% 
  tidyr::separate(files, into = c("Condition", "Replicate"), sep = "_(?=[^_]*$)") %>% 
  S4Vectors::DataFrame()
colnames(invitro.filtered) <- rownames(invitro.filtered@colData)
colnames(invitro.filtered@assays@data$counts) <- rownames(invitro.filtered@colData)
colnames(invitro.filtered@assays@data$abundance) <- rownames(invitro.filtered@colData)

invivo.filtered@colData <- invivo.filtered@colData %>% 
  as.data.frame %>% 
  dplyr::select(files) %>% 
  tidyr::separate(files, into = c("Condition", "Replicate"), sep = "_(?=[^_]*$)") %>% 
  S4Vectors::DataFrame()
colnames(invivo.filtered) <- rownames(invivo.filtered@colData)
colnames(invivo.filtered@assays@data$counts) <- rownames(invivo.filtered@colData)
colnames(invivo.filtered@assays@data$abundance) <- rownames(invivo.filtered@colData)
```

```{r}
invitro.filtered@colData %>% as.data.frame %>% dplyr::count(Condition)
```


```{r}
invivo.filtered@colData %>% as.data.frame %>% dplyr::count(Condition)
```

# DESeq Sample-Level normalization
```{r}
#DESeq objects
dds <- DESeq2::DESeqDataSetFromMatrix(countData=round(invitro.filtered@assays@data$counts),
                                      colData=invitro.filtered@colData,
                                      design=~Condition)
dds <- estimateSizeFactors(dds)

invitro.filtered@assays@data$rle <- counts(dds, normalized=TRUE)    
invitro.filtered@assays@data$rle_log2 <- log2(invitro.filtered@assays@data$rle + 1)
```

```{r}
#DESeq objects
dds <- DESeq2::DESeqDataSetFromMatrix(countData=round(invivo.filtered@assays@data$counts),
                                      colData=invivo.filtered@colData,
                                      design=~Condition)
dds <- estimateSizeFactors(dds)

invivo.filtered@assays@data$rle <- counts(dds, normalized=TRUE)    
invivo.filtered@assays@data$rle_log2 <- log2(invivo.filtered@assays@data$rle + 1)
```

# Save data
```{r}
if(do_save) {
  saveRDS(invitro.filtered,file=file.path(PATH,"data/yuhan_invitro.rds"))
  saveRDS(invivo.filtered,file=file.path(PATH,"data/yuhan_invivo.rds"))
} else {
  invitro.filtered <- readRDS(file.path(PATH,"data/yuhan_invitro.rds"))
  invivo.filtered <- readRDS(file.path(PATH,"data/yuhan_invivo.rds"))
}

```

# Visualizing Heteroskedasticity

## Mad vs Median {.tabset}

### invitro
```{r}
invitro.df <- data.frame(raw_median=matrixStats::rowMedians(invitro.filtered@assays@data$counts %>% as.matrix, na.rm=TRUE),
                 raw_mad=matrixStats::rowMads(invitro.filtered@assays@data$counts %>% as.matrix, na.rm=TRUE),
                 median=matrixStats::rowMedians(invitro.filtered@assays@data$rle_log2 %>% as.matrix, na.rm=TRUE),
                 mad=matrixStats::rowMads(invitro.filtered@assays@data$rle_log2 %>% as.matrix, na.rm=TRUE))

ggplot2::ggplot(invitro.df,aes(x=raw_median,y=raw_mad)) +
  geom_point() + 
  geom_smooth(mmethod = "loess") +
  labs(title = "Raw Mad vs Median")
```

### invivo
```{r}
invivo.df <- data.frame(raw_median=matrixStats::rowMedians(invivo.filtered@assays@data$counts %>% as.matrix, na.rm=TRUE),
                 raw_mad=matrixStats::rowMads(invivo.filtered@assays@data$counts %>% as.matrix, na.rm=TRUE),
                 median=matrixStats::rowMedians(invivo.filtered@assays@data$rle_log2 %>% as.matrix, na.rm=TRUE),
                 mad=matrixStats::rowMads(invivo.filtered@assays@data$rle_log2 %>% as.matrix, na.rm=TRUE))

ggplot2::ggplot(invivo.df,aes(x=raw_median,y=raw_mad)) +
  geom_point() + 
  geom_smooth(mmethod = "loess") +
  labs(title = "Raw Mad vs Median")
```


## Log+library normalized Mad vs Median {.tabset}

### invitro
```{r}
ggplot2::ggplot(invitro.df,aes(x=median,y=mad)) +
  geom_point() +
  geom_smooth(mmethod = "loess") + 
  labs(title = "Normalized Mad vs Median")
```

### invivo
```{r}
ggplot2::ggplot(invivo.df,aes(x=median,y=mad)) +
  geom_point() +
  geom_smooth(mmethod = "loess") + 
  labs(title = "Normalized Mad vs Median")
```

# Count Distributions 

## Unnormalized {.tabset}

### invitro
```{r}
invitro.filtered@assays@data$counts %>% melt %>% 
  ggplot(aes(x=variable, fill=variable, y = value)) + 
  geom_boxplot() +
  labs(title = "Sample-specific Distribution") +
  scale_x_discrete(guide = guide_axis(angle = 45)) 
```

### invivo 
```{r}
invivo.filtered@assays@data$counts %>% melt %>% 
  ggplot(aes(x=variable, fill=variable, y = value)) + 
  geom_boxplot() +
  labs(title = "Sample-specific Distribution") +
  scale_x_discrete(guide = guide_axis(angle = 45)) 
```


## DESeq Normalized {.tabset}

### invitro
```{r}
invitro.filtered@assays@data$rle_log2 %>% melt %>% 
  ggplot(aes(x=Var2, fill=Var2, y = value)) + 
  geom_boxplot() +
  labs(title = "Sample-specific Distribution") +
  scale_x_discrete(guide = guide_axis(angle = 45)) 
```

### invivo
```{r}
invivo.filtered@assays@data$rle_log2 %>% melt %>% 
  ggplot(aes(x=Var2, fill=Var2, y = value)) + 
  geom_boxplot() +
  labs(title = "Sample-specific Distribution") +
  scale_x_discrete(guide = guide_axis(angle = 45)) 
```


# Heatmaps {.tabset}

```{r}
hcopt <- function(d, HC=NULL, method = "ward.D", members = NULL){
  require("cba")
  if ( is.null(HC) ) {
    HC <- hclust(d,method=method,members=members)
  }
  #optimal leaf ordering
  ORD <- cba::order.optimal(d,merge=HC$merge)
  HC$merge <- ORD$merge
  HC$order <- ORD$order
  HC
}
```

## invitro 
```{r}
## Heatmap for the top 3000
dat_flt <- 
  invitro.filtered[order(matrixStats::rowMads(invitro.filtered@assays@data$rle_log2),decreasing = TRUE)[1:3000],]
## column annotation
annot <- invitro.filtered@colData %>% data.frame() %>% dplyr::select("Condition")
rownames(annot) <- colnames(invitro.filtered@assays@data$counts)
annot_col <- list(Condition=c(miR1453D = "lightblue",
                              miR1454D = "purple",
                              miR1513D = "orange",
                              miR1514D = "cyan",
                              sc_miRNA3D = "royalblue",
                              sc_miRNA4D = "gold"))
library(cba)
hc_row <- hcopt(as.dist(1-cor(t(dat_flt@assays@data$rle_log2))),method="ward.D")
hc_col <- hcopt(dist(t(dat_flt@assays@data$rle_log2)),method="ward.D")

pheatmap(dat_flt@assays@data$rle_log2,
         annotation_col = annot,
         annotation_colors = annot_col,
         cluster_rows=hc_row,
         cluster_cols=hc_col,
         show_rownames = FALSE,
         show_colnames = TRUE,
         scale = "row")
```

## invivo (IR/IS/Control)
```{r}
## Heatmap for the top 3000
dat_flt <- invivo.filtered[,1:14]
dat_flt <- dat_flt[order(matrixStats::rowMads(dat_flt@assays@data$rle_log2),decreasing = TRUE)[1:3000],]
## column annotation
annot <- dat_flt@colData %>% data.frame() %>% dplyr::select("Condition")
rownames(annot) <- colnames(dat_flt@assays@data$counts)
annot_col <- list(Condition=c(Control="gray",
                              IR="pink",
                              IS = "lightgreen"))
library(cba)
hc_row <- hcopt(as.dist(1-cor(t(dat_flt@assays@data$rle_log2))),method="ward.D")
hc_col <- hcopt(dist(t(dat_flt@assays@data$rle_log2)),method="ward.D")

pheatmap(dat_flt@assays@data$rle_log2,
         annotation_col = annot,
         annotation_colors = annot_col,
         cluster_rows=hc_row,
         cluster_cols=hc_col,
         show_rownames = FALSE,
         show_colnames = TRUE,
         scale = "row")
```

## invivo (miRNA145/scRNA)
```{r}
## Heatmap for the top 3000
dat_flt <- invivo.filtered[,15:21]
dat_flt <- dat_flt[order(matrixStats::rowMads(dat_flt@assays@data$rle_log2),decreasing = TRUE)[1:3000],]
## column annotation
annot <- dat_flt@colData %>% data.frame() %>% dplyr::select("Condition")
rownames(annot) <- colnames(dat_flt@assays@data$counts)
annot_col <- list(Condition=c(sc_miRNA="gray",
                              miR_145="pink"))
library(cba)
hc_row <- hcopt(as.dist(1-cor(t(dat_flt@assays@data$rle_log2))),method="ward.D")
hc_col <- hcopt(dist(t(dat_flt@assays@data$rle_log2)),method="ward.D")

pheatmap(dat_flt@assays@data$rle_log2,
         annotation_col = annot,
         annotation_colors = annot_col,
         cluster_rows=hc_row,
         cluster_cols=hc_col,
         show_rownames = FALSE,
         show_colnames = TRUE,
         scale = "row")
```