---
title: "Yuhan Primary Tumor RNA Seq - DESeq"
author: "Andrew Chen"
date:   "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: 'united'
    toc: true
    toc_depth: 1
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r}
library(Biobase)
library(SummarizedExperiment)
library(tidyverse)
library(ggplot2)
library(DESeq2)
library(reactable)
library(cba)
library(pheatmap)
library(vennr)
library(EnsDb.Hsapiens.v86)
library(AnnotationDbi)
library(EnsDb.Mmusculus.v79)
library(ggrepel)

PATH <- file.path(Sys.getenv("MLAB"),"projects/brcameta/projects/exosome/4t1_primary_tumor/")
DPATH <- file.path(Sys.getenv("CBM"),"otherStudies/RNAseq/")
```

# Load Data
```{r}
invitro.filtered <- readRDS(file.path(PATH,"data/yuhan_invitro.rds"))
invivo.filtered <- readRDS(file.path(PATH,"data/yuhan_invivo.rds"))

invivo_ir_is <- invivo.filtered[, 1:14]
invivo_mirna <- invivo.filtered[, 15:21]
```

# DESeq Result Tables {.tabset}

> DESeq2 performs an internal normalization where geometric mean is calculated for each gene across all samples. The counts for a gene in each sample is then divided by this mean. The median of these ratios in a sample is the size factor for that sample. This procedure corrects for library size and RNA composition bias, which can arise for example when only a small number of genes are very highly expressed in one experiment condition but not in the other. 

```{r}
invitro_dds <- DESeq2::DESeqDataSetFromMatrix(countData=round(invitro.filtered@assays@data$counts),
                                      colData=invitro.filtered@colData,
                                      design=~Condition)
invitro_dds <- estimateSizeFactors(invitro_dds)
invitro_dds$Condition <- relevel(invitro_dds$Condition, ref = "sc_miRNA4D")

invivo_iris_dds <- DESeq2::DESeqDataSetFromMatrix(countData=round(invivo_ir_is@assays@data$counts),
                                      colData=invivo_ir_is@colData,
                                      design=~Condition)
invivo_iris_dds <- estimateSizeFactors(invivo_iris_dds)
invivo_iris_dds$Condition <- relevel(invivo_iris_dds$Condition, ref = "Control")

invivo_mirna_dds <- DESeq2::DESeqDataSetFromMatrix(countData=round(invivo_mirna@assays@data$counts),
                                      colData=invivo_mirna@colData,
                                      design=~Condition)
invivo_mirna_dds <- estimateSizeFactors(invivo_mirna_dds)
invivo_mirna_dds$Condition <- relevel(invivo_mirna_dds$Condition, ref = "sc_miRNA")
```

> DESeq2 fits negative binomial generalized linear models for each gene and uses the Wald test for significance testing. Results were filtered so that only genes that had a |log fold change| > X and adjusted p-value < 0.05 (FDR) were retained.

```{r}
invitro_dds <- DESeq(invitro_dds)
invivo_iris_dds <- DESeq(invivo_iris_dds)
invivo_mirna_dds <- DESeq(invivo_mirna_dds)
```

## Invitro miR1514/scmiRNA

Summary of Different comparisons
```{r}
lfcs <- seq(0,1.0,0.1)
comp_df <- data.frame(lfcs = lfcs)
for (i in seq_along(lfcs)) {
  lfc <- lfcs[i]
  dds_res <- results(invitro_dds, alpha=0.05, lfcThreshold = lfc, contrast=c("Condition", "miR1514D", "sc_miRNA4D"))
  nPos <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange > lfc), na.rm=TRUE) 
  nNeg <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange < -lfc), na.rm=TRUE) 
  comp_df$nPos[i] = nPos
  comp_df$nNeg[i] = nNeg
}

comp_df
```


```{r}
dds_mir1514_scmiRNA_res <- results(invitro_dds, alpha=0.05, lfcThreshold = 0.1, contrast=c("Condition", "miR1514D", "sc_miRNA4D"))
dds_mir1514_scmiRNA_res_df <- dds_mir1514_scmiRNA_res[which(with(dds_mir1514_scmiRNA_res, padj<0.05)),] %>% as.data.frame
dds_mir1514_scmiRNA_res_df <- dds_mir1514_scmiRNA_res_df %>% dplyr::arrange(desc(log2FoldChange), padj)
symbols <- mapIds(EnsDb.Mmusculus.v79,
                  keys=rownames(dds_mir1514_scmiRNA_res_df),
                  column="SYMBOL",
                  keytype="GENEID")
dds_mir1514_scmiRNA_res_df$symbol <- symbols
write.csv(dds_mir1514_scmiRNA_res_df, file=file.path(PATH, "data/dds_invitro_mir1514D_scrna4D_diffanal.csv"), row.names = TRUE)
```

Table of Results
```{r}
reactable(dds_mir1514_scmiRNA_res_df, searchable = TRUE, filterable = TRUE)
```

## Invitro miR1454D/scmiRNA

Summary of Different comparisons
```{r}
lfcs <- seq(0,1.0,0.1)
comp_df <- data.frame(lfcs = lfcs)
for (i in seq_along(lfcs)) {
  lfc <- lfcs[i]
  dds_res <- results(invitro_dds, alpha=0.05, lfcThreshold = lfc, contrast=c("Condition", "miR1454D", "sc_miRNA4D"))
  nPos <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange > lfc), na.rm=TRUE) 
  nNeg <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange < -lfc), na.rm=TRUE) 
  comp_df$nPos[i] = nPos
  comp_df$nNeg[i] = nNeg
}

comp_df
```


```{r}
dds_mir1454_scmiRNA_res <- results(invitro_dds, alpha=0.05, lfcThreshold = 0.1, contrast=c("Condition", "miR1454D", "sc_miRNA4D"))
dds_mir1454_scmiRNA_res_df <- dds_mir1454_scmiRNA_res[which(with(dds_mir1454_scmiRNA_res, padj<0.05)),] %>% as.data.frame
dds_mir1454_scmiRNA_res_df <- dds_mir1454_scmiRNA_res_df %>% dplyr::arrange(desc(log2FoldChange), padj)
symbols <- mapIds(EnsDb.Mmusculus.v79,
                  keys=rownames(dds_mir1454_scmiRNA_res_df),
                  column="SYMBOL",
                  keytype="GENEID")
dds_mir1454_scmiRNA_res_df$symbol <- symbols
write.csv(dds_mir1454_scmiRNA_res_df, file=file.path(PATH, "data/dds_invitro_mir1454D_scrna4D_diffanal.csv"), row.names = TRUE)
```

Table of Results
```{r}
reactable(dds_mir1454_scmiRNA_res_df, searchable = TRUE, filterable = TRUE)
```

## Invitro scmiRNA3D/scmiRNA4D

Summary of Different comparisons
```{r}
lfcs <- seq(0,1.0,0.1)
comp_df <- data.frame(lfcs = lfcs)
for (i in seq_along(lfcs)) {
  lfc <- lfcs[i]
  dds_res <- results(invitro_dds, alpha=0.05, lfcThreshold = lfc, contrast=c("Condition", "sc_miRNA3D", "sc_miRNA4D"))
  nPos <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange > lfc), na.rm=TRUE) 
  nNeg <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange < -lfc), na.rm=TRUE) 
  comp_df$nPos[i] = nPos
  comp_df$nNeg[i] = nNeg
}

comp_df
```


```{r}
dds_3D_scmiRNA4D_res <- results(invitro_dds, alpha=0.05, lfcThreshold = 0.1, contrast=c("Condition", "sc_miRNA3D", "sc_miRNA4D"))
dds_3D_scmiRNA4D_res_df <- dds_3D_scmiRNA4D_res[which(with(dds_3D_scmiRNA4D_res, padj<0.05)),] %>% as.data.frame
dds_3D_scmiRNA4D_res_df <- dds_3D_scmiRNA4D_res_df %>% dplyr::arrange(desc(log2FoldChange), padj)
symbols <- mapIds(EnsDb.Mmusculus.v79,
                  keys=rownames(dds_3D_scmiRNA4D_res_df),
                  column="SYMBOL",
                  keytype="GENEID")
dds_3D_scmiRNA4D_res_df$symbol <- symbols
write.csv(dds_3D_scmiRNA4D_res_df, file=file.path(PATH, "data/dds_invitro_scrna3D_scrna4D_diffanal.csv"), row.names = TRUE)
```

Table of Results
```{r}
reactable(dds_3D_scmiRNA4D_res_df, searchable = TRUE, filterable = TRUE)
```

## Invitro miR1453D/miR1454D

```{r}
invitro_dds2 <- DESeq2::DESeqDataSetFromMatrix(countData=round(invitro.filtered@assays@data$counts),
                                      colData=invitro.filtered@colData,
                                      design=~Condition)
invitro_dds2 <- estimateSizeFactors(invitro_dds2)
invitro_dds2$Condition <- relevel(invitro_dds2$Condition, ref = "miR1454D")
invitro_dds2 <- DESeq(invitro_dds2)
```

Summary of Different comparisons
```{r}
lfcs <- seq(0,1.0,0.1)
comp_df <- data.frame(lfcs = lfcs)
for (i in seq_along(lfcs)) {
  lfc <- lfcs[i]
  dds_res <- results(invitro_dds2, alpha=0.05, lfcThreshold = lfc, contrast=c("Condition", "miR1453D", "miR1454D"))
  nPos <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange > lfc), na.rm=TRUE) 
  nNeg <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange < -lfc), na.rm=TRUE) 
  comp_df$nPos[i] = nPos
  comp_df$nNeg[i] = nNeg
}

comp_df
```


```{r}
dds_mir1453_1454 <- results(invitro_dds2, alpha=0.05, lfcThreshold = 0.3, contrast=c("Condition", "miR1453D", "miR1454D"))
dds_mir1453_1454_df <- dds_mir1453_1454[which(with(dds_mir1453_1454, padj<0.05)),] %>% as.data.frame
dds_mir1453_1454_df <- dds_mir1453_1454_df %>% dplyr::arrange(desc(log2FoldChange), padj)
symbols <- mapIds(EnsDb.Mmusculus.v79,
                  keys=rownames(dds_mir1453_1454_df),
                  column="SYMBOL",
                  keytype="GENEID")
dds_mir1453_1454_df$symbol <- symbols
write.csv(dds_mir1453_1454_df, file=file.path(PATH, "data/dds_invitro_mir1453_1454_diffanal.csv"), row.names = TRUE)
```

Table of Results
```{r}
reactable(dds_mir1453_1454_df, searchable = TRUE, filterable = TRUE)
```


## Invivo miRNA

Summary of Different comparisons
```{r}
lfcs <- seq(0,1.0,0.1)
comp_df <- data.frame(lfcs = lfcs)
for (i in seq_along(lfcs)) {
  lfc <- lfcs[i]
  dds_res <- results(invivo_mirna_dds, alpha=0.05, lfcThreshold = lfc, contrast=c("Condition", "miR_145", "sc_miRNA"))
  nPos <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange > lfc), na.rm=TRUE) 
  nNeg <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange < -lfc), na.rm=TRUE) 
  comp_df$nPos[i] = nPos
  comp_df$nNeg[i] = nNeg
}

comp_df
```


```{r}
dds_mir145_scmiRNA_res <- results(invivo_mirna_dds, alpha=0.05, lfcThreshold = 0, contrast=c("Condition", "miR_145", "sc_miRNA"))
dds_mir145_scmiRNA_res_df <- dds_mir145_scmiRNA_res[which(with(dds_mir145_scmiRNA_res, padj<0.05)),] %>% as.data.frame
dds_mir145_scmiRNA_res_df <- dds_mir145_scmiRNA_res_df %>% dplyr::arrange(desc(log2FoldChange), padj)
symbols <- mapIds(EnsDb.Mmusculus.v79,
                  keys=rownames(dds_mir145_scmiRNA_res_df),
                  column="SYMBOL",
                  keytype="GENEID")
dds_mir145_scmiRNA_res_df$symbol <- symbols
write.csv(dds_mir145_scmiRNA_res_df, file=file.path(PATH, "data/dds_invivo_mir145_scrna_diffanal.csv"), row.names = TRUE)
```

Table of Results
```{r}
reactable(dds_mir145_scmiRNA_res_df, searchable = TRUE, filterable = TRUE)
```

## Invivo IR/Control

Summary of Different comparisons
```{r}
lfcs <- seq(0,1.0,0.1)
comp_df <- data.frame(lfcs = lfcs)
for (i in seq_along(lfcs)) {
  lfc <- lfcs[i]
  dds_res <- results(invivo_iris_dds, alpha=0.05, lfcThreshold = lfc, contrast=c("Condition", "IR", "Control"))
  nPos <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange > lfc), na.rm=TRUE) 
  nNeg <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange < -lfc), na.rm=TRUE) 
  comp_df$nPos[i] = nPos
  comp_df$nNeg[i] = nNeg
}

comp_df
```


```{r}
dds_ir_control_res <- results(invivo_iris_dds, alpha=0.05, lfcThreshold = 0, contrast=c("Condition", "IR", "Control"))
dds_ir_control_res_df <- dds_ir_control_res[which(with(dds_ir_control_res, padj<0.05)),] %>% as.data.frame
dds_ir_control_res_df <- dds_ir_control_res_df %>% dplyr::arrange(desc(log2FoldChange), padj)
symbols <- mapIds(EnsDb.Mmusculus.v79,
                  keys=rownames(dds_ir_control_res_df),
                  column="SYMBOL",
                  keytype="GENEID")
dds_ir_control_res_df$symbol <- symbols
write.csv(dds_ir_control_res_df, file=file.path(PATH, "data/dds_invivo_ir_control_diffanal.csv"), row.names = TRUE)
```

Table of Results
```{r}
reactable(dds_ir_control_res_df, searchable = TRUE, filterable = TRUE)
```

## Invivo IR/IS

Summary of Different comparisons
```{r}
lfcs <- seq(0,1.0,0.1)
comp_df <- data.frame(lfcs = lfcs)
for (i in seq_along(lfcs)) {
  lfc <- lfcs[i]
  dds_res <- results(invivo_iris_dds, alpha=0.05, lfcThreshold = lfc, contrast=c("Condition", "IR", "IS"))
  nPos <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange > lfc), na.rm=TRUE) 
  nNeg <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange < -lfc), na.rm=TRUE) 
  comp_df$nPos[i] = nPos
  comp_df$nNeg[i] = nNeg
}

comp_df
```


```{r}
dds_iris_res <- results(invivo_iris_dds, alpha=0.05, lfcThreshold = 0, contrast=c("Condition", "IR", "IS"))
dds_iris_res_df <- dds_iris_res[which(with(dds_iris_res, padj<0.05)),] %>% as.data.frame
dds_iris_res_df <- dds_iris_res_df %>% dplyr::arrange(desc(log2FoldChange), padj)
symbols <- mapIds(EnsDb.Mmusculus.v79,
                  keys=rownames(dds_iris_res_df),
                  column="SYMBOL",
                  keytype="GENEID")
dds_iris_res_df$symbol <- symbols
write.csv(dds_iris_res_df, file=file.path(PATH, "data/dds_invivo_iris_diffanal.csv"), row.names = TRUE)
```

Table of Results
```{r}
reactable(dds_iris_res_df, searchable = TRUE, filterable = TRUE)
```

## Invivo IS/Control

Summary of Different comparisons
```{r}
lfcs <- seq(0,1.0,0.1)
comp_df <- data.frame(lfcs = lfcs)
for (i in seq_along(lfcs)) {
  lfc <- lfcs[i]
  dds_res <- results(invivo_iris_dds, alpha=0.05, lfcThreshold = lfc, contrast=c("Condition", "IS", "Control"))
  nPos <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange > lfc), na.rm=TRUE) 
  nNeg <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange < -lfc), na.rm=TRUE) 
  comp_df$nPos[i] = nPos
  comp_df$nNeg[i] = nNeg
}

comp_df
```


```{r}
dds_is_ctrl_res <- results(invivo_iris_dds, alpha=0.05, lfcThreshold = 0, contrast=c("Condition", "IS", "Control"))
dds_is_ctrl_res_df <- dds_is_ctrl_res[which(with(dds_is_ctrl_res, padj<0.05)),] %>% as.data.frame
dds_is_ctrl_res_df <- dds_is_ctrl_res_df %>% dplyr::arrange(desc(log2FoldChange), padj)
symbols <- mapIds(EnsDb.Mmusculus.v79,
                  keys=rownames(dds_is_ctrl_res_df),
                  column="SYMBOL",
                  keytype="GENEID")
dds_is_ctrl_res_df$symbol <- symbols
write.csv(dds_is_ctrl_res_df, file=file.path(PATH, "data/dds_invivo_is_ctrl_diffanal.csv"), row.names = TRUE)
```

Table of Results
```{r}
reactable(dds_is_ctrl_res_df, searchable = TRUE, filterable = TRUE)
```

# Signature Comparison
```{r}
# Preparing Signatures
invitro_1514_sc_up <- dds_mir1514_scmiRNA_res_df %>% dplyr::filter(log2FoldChange > 0) %>% rownames
invitro_1454_sc_up <- dds_mir1454_scmiRNA_res_df %>% dplyr::filter(log2FoldChange > 0) %>% rownames
invitro_scmiRNA3D_4D_up <- dds_3D_scmiRNA4D_res_df %>% dplyr::filter(log2FoldChange > 0 ) %>% rownames
invitro_miRNA1453_1454_up <- dds_mir1453_1454_df %>% dplyr::filter(log2FoldChange > 0 ) %>% rownames
invivo_miR_sc_up <- dds_mir145_scmiRNA_res_df %>% dplyr::filter(log2FoldChange > 0) %>% rownames
invivo_ir_is_up <- dds_iris_res_df %>% dplyr::filter(log2FoldChange > 0) %>% rownames
invivo_ir_ctrl_up <- dds_ir_control_res_df %>% dplyr::filter(log2FoldChange > 0) %>% rownames
invivo_is_ctrl_up <- dds_is_ctrl_res_df %>% dplyr::filter(log2FoldChange > 0) %>% rownames

invitro_1514_sc_dn <- dds_mir1514_scmiRNA_res_df %>% dplyr::filter(log2FoldChange < 0) %>% rownames
invitro_1454_sc_dn <- dds_mir1454_scmiRNA_res_df %>% dplyr::filter(log2FoldChange < 0) %>% rownames
invitro_scmiRNA3D_4D_dn <- dds_3D_scmiRNA4D_res_df %>% dplyr::filter(log2FoldChange < 0 ) %>% rownames
invitro_miRNA1453_1454_dn <- dds_mir1453_1454_df %>% dplyr::filter(log2FoldChange < 0 ) %>% rownames
invivo_miR_sc_dn <- dds_mir145_scmiRNA_res_df %>% dplyr::filter(log2FoldChange < 0) %>% rownames
invivo_ir_is_dn <- dds_iris_res_df %>% dplyr::filter(log2FoldChange < 0) %>% rownames
invivo_ir_ctrl_dn <- dds_ir_control_res_df %>% dplyr::filter(log2FoldChange < 0) %>% rownames
invivo_is_ctrl_dn <- dds_is_ctrl_res_df %>% dplyr::filter(log2FoldChange < 0) %>% rownames

# Combining labels
all_sigs <- list(mir1514_sc_up=invitro_1514_sc_up, mir1454_sc_up=invitro_1454_sc_up, miR_sc_up=invivo_miR_sc_up, ir_is_up=invivo_ir_is_up, ir_ctrl_up=invivo_ir_ctrl_up, is_ctrl_up = invivo_is_ctrl_up,
                 mir1514_sc_dn=invitro_1514_sc_dn, mir1454_sc_dn=invitro_1454_sc_dn, miR_sc_dn=invivo_miR_sc_dn, ir_is_dn=invivo_ir_is_dn, ir_ctrl_dn=invivo_ir_ctrl_dn, is_ctrl_dn = invivo_is_ctrl_dn, miR1453_1454_up = invitro_miRNA1453_1454_up, miR1453_1454_dn = invitro_miRNA1453_1454_dn, sc3D_4D_up = invitro_scmiRNA3D_4D_up, sc3D_4D_dn = invitro_scmiRNA3D_4D_dn)
```

```{r}
vennr(all_sigs)
```


# Heatmaps {.tabset}

```{r}
hcopt <- function(d, HC=NULL, method = "ward.D", members = NULL){
  require("cba")
  if ( is.null(HC) ) {
    HC <- hclust(d,method=method,members=members)
  }
  #optimal leaf ordering
  ORD <- cba::order.optimal(d,merge=HC$merge)
  HC$merge <- ORD$merge
  HC$order <- ORD$order
  HC
}
```

## Invitro miR1514/scmiRNA
```{r}
## Heatmap for the top 3000
dat_flt <- 
  invitro.filtered[c(invitro_1514_sc_up, invitro_1514_sc_dn),c(7:9,14:16)]
## column annotation
annot <- invitro.filtered@colData %>% data.frame() %>% dplyr::select("Condition")
rownames(annot) <- colnames(invitro.filtered@assays@data$counts)
annot_col <- list(Condition=c(sc_miRNA4D="gray",miR1514D="pink"))
library(cba)
hc_row <- hcopt(as.dist(1-cor(t(dat_flt@assays@data$rle_log2))),method="ward.D")
hc_col <- hcopt(dist(t(dat_flt@assays@data$rle_log2)),method="ward.D")

pheatmap(dat_flt@assays@data$rle_log2,
         annotation_col = annot,
         annotation_colors = annot_col,
         cluster_rows=hc_row,
         cluster_cols=hc_col,
         show_rownames = FALSE,
         show_colnames = TRUE,
         scale = "row")
```

## Invitro miR1454/scmiRNA
```{r}
## Heatmap for the top 3000
dat_flt <- 
  invitro.filtered[c(invitro_1454_sc_up, invitro_1454_sc_dn),c(3,10,14:16)]
## column annotation
annot <- invitro.filtered@colData %>% data.frame() %>% dplyr::select("Condition")
rownames(annot) <- colnames(invitro.filtered@assays@data$counts)
annot_col <- list(Condition=c(sc_miRNA4D="gray",miR1454D="pink"))
library(cba)
hc_row <- hcopt(as.dist(1-cor(t(dat_flt@assays@data$rle_log2))),method="ward.D")
hc_col <- hcopt(dist(t(dat_flt@assays@data$rle_log2)),method="ward.D")

pheatmap(dat_flt@assays@data$rle_log2,
         annotation_col = annot,
         annotation_colors = annot_col,
         cluster_rows=hc_row,
         cluster_cols=hc_col,
         show_rownames = FALSE,
         show_colnames = TRUE,
         scale = "row")
```

## Invitro scmiRNA3D/scmiRNA4D
```{r}
## Heatmap for the top 3000
dat_flt <- 
  invitro.filtered[c(invitro_scmiRNA3D_4D_up, invitro_scmiRNA3D_4D_dn),c(11:16)]
## column annotation
annot <- invitro.filtered@colData %>% data.frame() %>% dplyr::select("Condition")
rownames(annot) <- colnames(invitro.filtered@assays@data$counts)
annot_col <- list(Condition=c(sc_miRNA4D="gray",sc_miRNA3D="pink"))
library(cba)
hc_row <- hcopt(as.dist(1-cor(t(dat_flt@assays@data$rle_log2))),method="ward.D")
hc_col <- hcopt(dist(t(dat_flt@assays@data$rle_log2)),method="ward.D")

pheatmap(dat_flt@assays@data$rle_log2,
         annotation_col = annot,
         annotation_colors = annot_col,
         cluster_rows=hc_row,
         cluster_cols=hc_col,
         show_rownames = FALSE,
         show_colnames = TRUE,
         scale = "row")
```

## Invitro miR1453D/miR1454D
```{r}
## Heatmap for the top 3000
dat_flt <- 
  invitro.filtered[c(invitro_miRNA1453_1454_up, invitro_miRNA1453_1454_dn),c(1:2,3,10)]
## column annotation
annot <- invitro.filtered@colData %>% data.frame() %>% dplyr::select("Condition")
rownames(annot) <- colnames(invitro.filtered@assays@data$counts)
annot_col <- list(Condition=c(miR1454D="gray",miR1453D="pink"))
library(cba)
hc_row <- hcopt(as.dist(1-cor(t(dat_flt@assays@data$rle_log2))),method="ward.D")
hc_col <- hcopt(dist(t(dat_flt@assays@data$rle_log2)),method="ward.D")

pheatmap(dat_flt@assays@data$rle_log2,
         annotation_col = annot,
         annotation_colors = annot_col,
         cluster_rows=hc_row,
         cluster_cols=hc_col,
         show_rownames = FALSE,
         show_colnames = TRUE,
         scale = "row")
```
## Invivo miRNA
```{r}
## Heatmap for the top 3000
dat_flt <- 
  invivo.filtered[c(invivo_miR_sc_up, invivo_miR_sc_dn),15:21]
## column annotation
annot <- invivo.filtered@colData %>% data.frame() %>% dplyr::select("Condition")
rownames(annot) <- colnames(invivo.filtered@assays@data$counts)
annot_col <- list(Condition=c(sc_miRNA="gray",miR_145="pink"))
library(cba)
hc_row <- hcopt(as.dist(1-cor(t(dat_flt@assays@data$rle_log2))),method="ward.D")
hc_col <- hcopt(dist(t(dat_flt@assays@data$rle_log2)),method="ward.D")

pheatmap(dat_flt@assays@data$rle_log2,
         annotation_col = annot,
         annotation_colors = annot_col,
         cluster_rows=hc_row,
         cluster_cols=hc_col,
         show_rownames = FALSE,
         show_colnames = TRUE,
         scale = "row")
```

## Invivo IR/IS
```{r}
## Heatmap for the top 3000
dat_flt <- 
  invivo.filtered[c(invivo_ir_is_up, invivo_ir_is_dn),5:14]
rownames(dat_flt@assays@data$rle_log2) <- rowData(dat_flt)$gene_name
## column annotation
annot <- invivo.filtered@colData %>% data.frame() %>% dplyr::select("Condition")
rownames(annot) <- colnames(invivo.filtered@assays@data$counts)
annot_col <- list(Condition=c(IS="gray",IR="pink"))
library(cba)
hc_row <- hcopt(as.dist(1-cor(t(dat_flt@assays@data$rle_log2))),method="ward.D")
hc_col <- hcopt(dist(t(dat_flt@assays@data$rle_log2)),method="ward.D")

pheatmap(dat_flt@assays@data$rle_log2,
         annotation_col = annot,
         annotation_colors = annot_col,
         cluster_rows=hc_row,
         cluster_cols=hc_col,
         show_rownames = TRUE,
         show_colnames = TRUE,
         scale = "row")
```

## Invivo IR/Control
```{r}
## Heatmap for the top 3000
dat_flt <- 
  invivo.filtered[c(invivo_ir_ctrl_up, invivo_ir_ctrl_dn), 1:9]
## column annotation
annot <- invivo.filtered@colData %>% data.frame() %>% dplyr::select("Condition")
rownames(annot) <- colnames(invivo.filtered@assays@data$counts)
annot_col <- list(Condition=c(Control="gray",IR="pink"))
library(cba)
hc_row <- hcopt(as.dist(1-cor(t(dat_flt@assays@data$rle_log2))),method="ward.D")
hc_col <- hcopt(dist(t(dat_flt@assays@data$rle_log2)),method="ward.D")

pheatmap(dat_flt@assays@data$rle_log2,
         annotation_col = annot,
         annotation_colors = annot_col,
         cluster_rows=hc_row,
         cluster_cols=hc_col,
         show_rownames = FALSE,
         show_colnames = TRUE,
         scale = "row")
```

## Invivo IS/Control
```{r}
## Heatmap for the top 3000
dat_flt <- 
  invivo.filtered[c(invivo_is_ctrl_up, invivo_is_ctrl_dn), c(1:4, 10:14)]
rownames(dat_flt@assays@data$rle_log2) <- rowData(dat_flt)$gene_name
## column annotation
annot <- invivo.filtered@colData %>% data.frame() %>% dplyr::select("Condition")
rownames(annot) <- colnames(invivo.filtered@assays@data$counts)
annot_col <- list(Condition=c(Control="gray",IS="pink"))
library(cba)
hc_row <- hcopt(as.dist(1-cor(t(dat_flt@assays@data$rle_log2))),method="ward.D")
hc_col <- hcopt(dist(t(dat_flt@assays@data$rle_log2)),method="ward.D")

pheatmap(dat_flt@assays@data$rle_log2,
         annotation_col = annot,
         annotation_colors = annot_col,
         cluster_rows=hc_row,
         cluster_cols=hc_col,
         show_rownames = TRUE,
         show_colnames = TRUE,
         scale = "row")
```

# Volcano Plot

```{r}
dds_iris_res
symbols <- mapIds(EnsDb.Mmusculus.v79,
                  keys=rownames(dds_iris_res),
                  column="SYMBOL",
                  keytype="GENEID")
dds_iris_res$symbol <- symbols
```

```{r}
dds_iris_res %>% 
  as.data.frame %>% 
  dplyr::filter(!is.na(symbol) & !is.na(padj)) %>%
  dplyr::mutate(
    Significant = if_else(
      padj < 0.05,
      TRUE,
      FALSE)) %>%
  dplyr::mutate(
    label = if_else(
      Significant,
      symbol,
      ""
    )
  ) %>%
  ggplot(aes(x = log2FoldChange, y = -log10(padj), color = Significant, label = label)) +
  geom_point() +
  geom_text_repel(min.segment.length = 0,
                  nudge_x = 0.25,
                  nudge_y = 0.25,
                  segment.size = 0.2) +
  theme_minimal() +
  labs(title = "IR/IS Volcano Plot",
       x = "log2 Fold Change",
       y = "-log10 p-value")
```




