---
title: "4T1 Mets RNA Seq - Differential Analysis"
author: "Andrew Chen"
date:   "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: 'united'
    toc: true
    toc_depth: 1
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r}
library(Biobase)
library(SummarizedExperiment)
library(tidyverse)
library(ggplot2)
library(DESeq2)
library(reactable)
library(cba)
library(pheatmap)
library(vennr)

PATH <- file.path(Sys.getenv("MLAB"),"projects/brcameta/exosome/4t1_brca_brain_mets/")
DPATH <- file.path(Sys.getenv("CBM"),"otherStudies/RNAseq/2022-06-03-DenisExosomeBrCaBrainMets/")
```

# Load Data
```{r}
dat <- readRDS(file.path(PATH,"data/4T1_mets_sExp.rds"))
dat_pool <- readRDS(file.path(PATH,"data/4T1_mets_control_IS_pool_sExp.rds"))
```

# DESeq Result Tables {.tabset}

> DESeq2 performs an internal normalization where geometric mean is calculated for each gene across all samples. The counts for a gene in each sample is then divided by this mean. The median of these ratios in a sample is the size factor for that sample. This procedure corrects for library size and RNA composition bias, which can arise for example when only a small number of genes are very highly expressed in one experiment condition but not in the other. 

```{r}
dds <- DESeq2::DESeqDataSetFromMatrix(countData=round(assays(dat)$counts),
                                      colData=dat@colData,
                                      design=~condition)
```

> DESeq2 fits negative binomial generalized linear models for each gene and uses the Wald test for significance testing. Results were filtered so that only genes that had a |log fold change| > 0.5 and adjusted p-value < 0.05 (FDR) were retained.

```{r}
# Map EnsemblID to GeneID
ensembl_symbol <- function(ensembl_id, sExp) {
  return(unlist(rowData(sExp)[ensembl_id,"mgi_symbol"]))
}
symbol_ensembl <- function(gene_id, sExp) {
  return(rownames(sExp)[rowData(sExp)$mgi_symbol == gene_id])
}
```

## IR IS
```{r}
dds <- DESeq(dds)
```

Summary of Different comparisons
```{r}
lfcs <- seq(0,0.5,0.1)
comp_df <- data.frame(lfcs = lfcs)
for (i in seq_along(lfcs)) {
  lfc <- lfcs[i]
  dds_res <- results(dds, alpha=0.05, lfcThreshold = lfc, contrast=c("condition", "IR", "IS"))
  nPos <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange > lfc), na.rm=TRUE) 
  nNeg <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange < -lfc), na.rm=TRUE) 
  comp_df$nPos[i] = nPos
  comp_df$nNeg[i] = nNeg
}

comp_df
```


```{r}
dds_ir_is_res <- results(dds, alpha=0.05, lfcThreshold = 0.1, contrast=c("condition", "IR", "IS"))
dds_ir_is_res_df <- dds_ir_is_res[which(with(dds_ir_is_res, padj<0.05 & (log2FoldChange <= -0.1 | log2FoldChange >= 0.1))),] %>% as.data.frame
dds_ir_is_res_df_symbol <- unlist(lapply(rownames(dds_ir_is_res_df), ensembl_symbol, sExp=dat))
dds_ir_is_res_df <- dds_ir_is_res_df[dds_ir_is_res_df_symbol != "",]
rownames(dds_ir_is_res_df) <- dds_ir_is_res_df_symbol[dds_ir_is_res_df_symbol != ""]
write.csv(dds_ir_is_res_df[order(dds_ir_is_res_df$log2FoldChange),], file=file.path(PATH, "data/dds_ir_is_diffanal.csv"))
```

Table of Results
```{r}
reactable(dds_ir_is_res_df, searchable = TRUE)
```

## IR Control 
```{r}
lfcs <- seq(0,1,0.1)
comp_df <- data.frame(lfcs = lfcs)
for (i in seq_along(lfcs)) {
  lfc <- lfcs[i]
  dds_res <- results(dds, alpha=0.05, lfcThreshold = lfc, contrast=c("condition", "IR", "C"))
  nPos <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange > lfc), na.rm=TRUE) 
  nNeg <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange < -lfc), na.rm=TRUE) 
  comp_df$nPos[i] = nPos
  comp_df$nNeg[i] = nNeg
}

comp_df
```


```{r}
dds_ir_c_res <- results(dds, alpha=0.05, lfcThreshold = 0.4, contrast=c("condition", "IR", "C"))
summary(dds_ir_c_res)

dds_ir_c_res_df <- dds_ir_c_res[which(with(dds_ir_c_res, padj<0.05 & (log2FoldChange < -0.4 | log2FoldChange > 0.4))),] %>% as.data.frame
dds_ir_c_res_df_symbol <- unlist(lapply(rownames(dds_ir_c_res_df), ensembl_symbol, sExp=dat))
dds_ir_c_res_df <- dds_ir_c_res_df[dds_ir_c_res_df_symbol != "",]
rownames(dds_ir_c_res_df) <- dds_ir_c_res_df_symbol[dds_ir_c_res_df_symbol != ""]
write.csv(dds_ir_c_res_df[order(dds_ir_c_res_df$log2FoldChange),], file=file.path(PATH, "data/dds_IR_C_diffanal.csv"))
```

Table of Results
```{r}
reactable(dds_ir_c_res_df, searchable = TRUE)
```


## Control IS
```{r}
lfcs <- seq(0,0.5,0.1)
comp_df <- data.frame(lfcs = lfcs)
for (i in seq_along(lfcs)) {
  lfc <- lfcs[i]
  dds_res <- results(dds, alpha=0.05, lfcThreshold = lfc, contrast=c("condition", "C", "IS"))
  nPos <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange > lfc), na.rm=TRUE) 
  nNeg <- sum((dds_res$padj < 0.05) & (dds_res$log2FoldChange < -lfc), na.rm=TRUE) 
  comp_df$nPos[i] = nPos
  comp_df$nNeg[i] = nNeg
}

comp_df
```

```{r}
dds_c_is <- results(dds, alpha=0.05, lfcThreshold = 0.3, contrast=c("condition", "C", "IS"))
summary(dds_c_is)

dds_c_is_df <- as.data.frame(dds_c_is[which(dds_c_is$padj < 0.05, dds_c_is$log2FoldChange > 0.3),])
dds_c_is_df_symbol <- unlist(lapply(rownames(dds_c_is_df), ensembl_symbol, sExp=dat_pool))
dds_c_is_df <- dds_c_is_df[dds_c_is_df_symbol != "",]
rownames(dds_c_is_df) <- dds_c_is_df_symbol[dds_c_is_df_symbol != ""]
write.csv(dds_c_is_df[order(dds_c_is_df$log2FoldChange),], file=file.path(PATH,"data/dds_c_isontrol_pool_diffanal.csv"))
```

Table of Results
```{r}
reactable(dds_c_is_df, searchable = TRUE)
```

> No significantly upregulated genes if IS and C are pooled given the same significance and logFC thresholds

# Signature Comparison
```{r}
# Preparing Signatures
ir_c_up_signatures <- dds_ir_c_res_df %>% dplyr::filter(log2FoldChange > 0) %>% rownames
c_is_up_signatures <- dds_c_is_df %>% dplyr::filter(log2FoldChange > 0) %>% rownames
ir_is_up_signatures <- dds_ir_is_res_df %>% dplyr::filter(log2FoldChange > 0) %>% rownames

ir_c_down_signatures <- dds_ir_c_res_df %>% dplyr::filter(log2FoldChange < 0) %>% rownames
c_is_down_signatures <- dds_c_is_df %>% dplyr::filter(log2FoldChange < 0) %>% rownames
ir_is_down_signatures <- dds_ir_is_res_df %>% dplyr::filter(log2FoldChange < 0) %>% rownames

# Combining labels
all_sigs <- list(ir_c_up=ir_c_up_signatures, c_is_up=c_is_up_signatures, ir_is_up=ir_is_up_signatures,
                 ir_c_down=ir_c_down_signatures, c_is_down=c_is_down_signatures,
                 ir_is_down=ir_is_down_signatures)
ensembl_sigs <- lapply(all_sigs, function(x) unlist(lapply(x, symbol_ensembl, sExp=dat)))
```

```{r}
vennr(all_sigs)
```


# Heatmaps {.tabset}

```{r}
hcopt <- function(d, HC=NULL, method = "ward.D", members = NULL){
  require("cba")
  if ( is.null(HC) ) {
    HC <- hclust(d,method=method,members=members)
  }
  #optimal leaf ordering
  ORD <- cba::order.optimal(d,merge=HC$merge)
  HC$merge <- ORD$merge
  HC$order <- ORD$order
  HC
}
```

```{r}
unique_union_de_genes <- unique(unlist(ensembl_sigs))
dat_flt <- dat[unique_union_de_genes,]

dds_annot_col <- colData(dat_flt) %>% data.frame %>% dplyr::select(condition)

dds_annot_row <- data.frame(
  ir_c = dds_ir_c_res %>% as.data.frame() %>% dplyr::filter(row.names(dds_ir_c_res) %in% unique_union_de_genes) %>%
    dplyr::transmute(ir_c=ifelse(log2FoldChange>0,"up","down")),
  c_is = dds_c_is %>% as.data.frame() %>% dplyr::filter(row.names(dds_c_is) %in% unique_union_de_genes) %>%
    dplyr::transmute(c_is=ifelse(log2FoldChange>0,"up","down")),
  ir_is = dds_ir_is_res %>% as.data.frame() %>% dplyr::filter(row.names(dds_ir_is_res) %in% unique_union_de_genes) %>%
    dplyr::transmute(ir_is=ifelse(log2FoldChange>0,"up","down"))
)

dds_annot_colors <- list(condition=c(C="gray",IR="pink", IS="orange"),
                     ir_c=c(up="red",down="blue"),
                     c_is=c(up="red",down="blue"),
                     ir_is=c(up="red",down="blue"))
  
## use Euclidean distance for columns/samples
## use ward as agglomeration rule
dds.hc01.col <- hcopt(dist(t(assays(dat_flt)$rle_log2)),method="ward.D2")

## use 1-correlation as distance for for rows/genes
## use ward as agglomeration rule
dds.hc01.row <- hcopt(as.dist(1-cor(t(assays(dat_flt)$rle_log2))),method="ward.D2")
```


```{r}
pheatmap(assays(dat_flt)$rle_log2,
         annotation_col = dds_annot_col,
         annotation_row = dds_annot_row,
         annotation_colors = dds_annot_colors,
         cluster_rows=dds.hc01.row,
         cluster_cols=dds.hc01.col,
         show_rownames = FALSE,
         show_colnames = TRUE,
         border_color = NA,
         scale = "row")
```

> Among DE genes, IS is more closely related to Control. 


# Saving Signatures
```{r}
deseq_list <- list(ir_is = dds_ir_is_res, ir_c = dds_ir_c_res, c_is = dds_c_is)
saveRDS(deseq_list,file=file.path(PATH,"data/deseq_list.rds"))
saveRDS(all_sigs,file=file.path(PATH,"data/signatures_symbol.rds"))
```

